Unified data structures for solving optimally a set of interrelated computational geometry problems

### 2019-12-23 SoCG EA/article-SoCG-EA.tex
SoCG EA, SoCG, acm siggraph + restructuring

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{diagbox}
\usepackage{amsthm}
\usepackage[formats]{listings}
\lstset{
	language=Java,
	basicstyle=\small,
	breaklines,
}
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\theoremstyle{plane}
\newtheorem{theorem}{Theorem}[section]

@@

\title{Unified data structures for solving optimally a set of interrelated computational geometry problems*}
\author{
\IEEEauthorblockN{1\textsuperscript{st} Vasyl Tereshchenko}
\IEEEauthorblockA{\textit{Faculty of computer science and cybernetics} \\
\textit{Taras Shevchenko National University of Kyiv}\\
Kyiv , Ukraine \\
vtereshch@gmail.com}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Semen Chudakov}
\IEEEauthorblockA{\textit{Faculty of computer science and cybernetics} \\
\textit{Taras Shevchenko National University of Kyiv}\\
Kyiv , Ukraine \\
semen.chudakov7@gmail.com}
}
\maketitle

\begin{abstract}
		The paper is devoted to the development and  of an efficient algorithmic model for solving a set of interrelated computational geometry problems. To do this, a unified algorithmic environment with unified data structures is created, which allows to implement complex use cases efficiently with respect to computational resources. We build the environment on the basis of the ``divide and conquer'' strategy. 
		Once a convex hull is key to a set of computational geometry problems, we offer a concatenable queue data structure to maintain it. The data structure is implemented in a form of a binary tree. This allows to perform operations needed in algorithm for a set of problems in $O(\log n)$ time. Furthermore we offer a way to execute the algorithms both sequentially and in parallel.
		In the future the algorithmic environment can be improved to support other computational models with similar properties for solving problems. As an example, the Voronoi diagram or the Delaunay triangulation can be considered.
\end{abstract}

\begin{IEEEkeywords}
Algorithmic tools,   Computational geometry, Interrelated problems set, Unified algorithmic environment, Concatenable queue
\end{IEEEkeywords}

\section{Introduction}

@@

\section{Unified algorithmic environment}
\subsection{Algorithms stages}

@@

		\begin{equation}
			\lambda(a_1, a_2)=\frac{y(a_2)-y(a_1)}{x(a_2)-x(a_1)}
		\end{equation}
		According to this definition, the possible cases for points $a_1,a_2,a_3$ are given in Table \ref{table:3points}:

		\begin{table}[htbp]
			\caption{3-points base cases}
			\label{table:3points}
			\begin{center}
				\begin{tabular}{|c|c|c|c|}
					\hline
					\textbf{$\lambda(a_1, a_2) > \lambda(a_2, a_3)$} & \textbf{$y(a_1) < y(a_2)$} & upper part & lower part\\
					\hline
					false & false & $a_1, a_3$ & $a_2$ \\
					\hline
					false & true & $a_1, a_3$ & $a_2$ \\
					\hline
					true & false & $a_2, a_3$ & $a_1$ \\
					\hline
					true & true & $a_1, a_2$ & $a_3$ \\
					\hline
				\end{tabular} 
			\end{center}
		\end{table}

@@

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.4\textwidth, height=0.2\textheight]{ch_union}}
			\caption{Merging step of two hulls}
			\label{fig:ch_union}
		\end{figure}

@@

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.4\textwidth, height=0.2\textheight]{incorect_search}}
			\caption{Example of an incorrect position of the utmost left in the left sub-hull}
			\label{fig:incorect_search}
		\end{figure}

@@

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.4\textwidth, height=0.2\textheight]{incorect_edge_points}}
			\caption{Example of a convex hull for a wrong position of the utmost left points of the left sub-hull}
			\label{fig:incorect_edge_points}
		\end{figure}

@@

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.4\textwidth, height=0.2\textheight]{incorect_lower_subhull}}
			\caption{An example of a non-integral hull after merging along the reference lines}
			\label{fig:incorect_lower_subhull}
		\end{figure}

@@

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.4\textwidth, height=0.2\textheight]{correct_convex_hull}}
			\caption{Correctly constructed convex hull}
			\label{fig:correct_convex_hull}
		\end{figure}

@@

		\begin{lstlisting}[caption={Algorithm model based on the ``divide-and-conquer'' principle},label={lst:algmodel},captionpos=b]
interface DaCAlgorithm[IT, OT]:

  boolean isBaseCase(IT input)
  int inputSize(IT input)

@@

  OT merge(OT first, OT second)
  Pair[IT, IT] divide(IT input)

		\end{lstlisting}

\subsection{Sequential and parallel execution}

		Although this model very accurately describes the class of algorithms, it does not make it possible to solve the problem directly by having input data. This in fact allows to divide implementation of the algorithm from how it is executed. Next the principles of sequential and parallel execution are discussed.

		When executing sequentially an algorithm, its individual sub-problems are computed one by one. We first check if current input is a base case and if so we can directly compute it by calling $solveBaseCase$ procedure. Otherwise input is split with $divide$ and obtained sub-problem are solved sequentially. Finally obtained results are merged with $merge$ procedure.

		In parallel execution, we take into account that the individual sub-problems can be calculated independently, which significantly speeds up the execution of the algorithm.

		To construct the concurrent execution algorithm, we use the following parallel computation abstraction $computeInParallel(function1, function2)$. which runs the functions $function1$ and $function2$ simultaneously. We use it to solve sub-problem obtained after splitting a given input. Other than that parallel version is identical to the sequential one.

		Practically, from the implementation standpoint performance of parallel execution was improved by introducing a limit on the size of sub-tasks that can be calculated in parallel. This allowed to put a threshold on the amount of work for one thread.

\section{Implementation details}
\subsection{Concatenable queue}

		As shown in \cite{overmars}, the concatenable queue is the key data structure for the algorithm described above and is therefore the basis for the UAEM. Now we will focus on how to efficiently implement it in our algorithmic environment.

		Concatenable queue is an Abstract Data Type, that supports following operations:

		\begin{itemize}
			\item
			ADD\_ELEMENT();
			\item
			REMOVE\_ELEMENT();
			\item
			GET\_MINIMUM();
			\item
			CONTAINS();
			\item
			SPLIT();
			\item
			MERGE().
		\end{itemize}

		By default the elements in a concatenable queue are kept i a certain predefined order \cite[pp..~155-157]{aho}.

		In this article the concatenable queue is implemented as a binary  $B+$ tree. Its vertices are divided into non-leaf and leaf ones. The leaf vertices contain all data kept in a tree. Every vertex has a pointer to its left and right child. For the leaf vertices those pointers point to the left and right neighboring leaf vertices or $null$ if the vertex is utmost in the tree. 
		Additionally every vertex keeps a pointer to a vertex with the largest element in its left sub-tree, which  allows to perform binary search \cite[pp..~155-157]{aho}. The height of each vertex is kept for the balancing during the split and merge operations. It is measured as a maximum amount of steps it is possible to do in order to reach a leaf vertex.

		From now we will go into details on how this data structure is implemented.

		The contains operations is pretty straightforward and uses binary search over the tree so its complexity is $O(\log n)$, where $n$ hereafter denotes the numbers of vertices in the queue.

		Algorithm of inserting an element in a queue looks like as follows. First, the position of a new vertex is searched. Then the connections between adjacent leaf vertices are broken to insert a new one. Going back a new non-leaf vertex is created - the parent for the new element and one of its neighbor. The algorithm is formally described on the Listing \ref{lst:insert-alg}.

		\begin{figure}[htbp]
			\begin{lstlisting}[caption={Queue insertion algorithm},label={lst:insert-alg},captionpos=b]
Node insert(Node node, int e):
  Node result = nil
  if e <= node.leftSubtreeMax.data: 
  	if node.isLeaf:
  	  if e == node.data:
        node.data = e
  	  else:
        Node createdLeaf = createLeafBetween(e, node.left, node)
        result = Node(createdLeaf, createdLeaf, node)
    else:
      node.left = insert(node.left, e)
  else:
    if node.isLeaf: 
      Node createdLeaf = createLeafBetween(e, node, node.right)
      result = Node(node, node, createdLeaf)
    else: 
      node.right = insert(node.right, e)

  if result == nil:
  	result = node

  updateHeight(result)
  return result
			\end{lstlisting}
		\end{figure}

		Here the \textit{updateHeight} subroutine updates the height value for a given vertex, the \textit{createLeafBetween} subroutine breaks the connection between adjacent vertices ot insert a new one.

		On the first step we find out if $leftSubtreeMax$ point to an element with greater value than the value to be inserted $e$. If so, then, if current element is a leaf, a new element is created between the current element and its left neighbor. Otherwise the search proceeds on the left sub-tree of the current element. The case, when $leftSubtreeMax$ is smaller than the value $e$ is analogous. The procedure ends with updating the height on a newly created element. The element is returned as its result value.

		Since on every step we perform a constant amount of work, the complexity of the procedure is $O(h)=O(\log n)$, where $h$ hereafter denotes the height of the tree. The operation of removing an element from the queue is performed analogously.

		The split operation is a bit more complex. As an input the procedure takes current element and value based on which the split is performed. As a result two independent queues are formed. The value, by which the split has been performed, belongs to the left sub-queue. The procedure is formally described on the Listing \ref{lst:split-alg}.

		\begin{lstlisting}[caption={Queue split algorithm},label={lst:split-alg},captionpos=b]
split(Node node, int e, ConcatenableQueue leftQueue, ConcatenableQueue rightQueue): 
  if node.isLeaf
    leftQueue.root = node
    leftQueue.maxNode = node	
    rightQueue.minNode = node.right	
    cut(node)
  else:
    if e == node.leftSubtreeMax.data:
      leftQueue.root = node.left
      leftQueue.maxNode = node.leftSubtreeMax

      rightQueue.root = node.right
      rightQueue.minNode = node.leftSubtreeMax.right

      cut(node.leftSubtreeMax)
    else if e < node.leftSubtreeMax.data:
      split(node.left, e, leftQueue, rightQueue)
      rightQueue.root = concatenateNodes(rightQueue.root, node.right, node.leftSubtreeMax)
    else:
      split(node.right, e, leftQueue, rightQueue)
      leftQueue.root = concatenateNodes(node.left,  leftQueue.root, node.leftSubtreeMax)
		\end{lstlisting}

		Here the $concatenateNodes$ procedure is used. It performs concatenation of two arbitrary elements and uses theirs heights to balance the resulting queue. The $cut$ procedure breaks connection between two adjacent leaf elements in a queue and therefore is trivial. On the first step in the $split$ operation we check, is the current element is a leaf. Is so, its connection are broken and the value of $maxNode$ is updated for the left queue as well as the value of $minNode$ for the right queue. If element is not a leaf, then the procedure continues of either left or right sub-tree. Here a special corner cases is considered, where $leftSubtreeMax$ contain the dividing value. Then the analogous action to the usual search are performed.

		The algorithm of the $concatenateNodes$ procedure is described on the Listing \ref{lst:concatenate-als}.

		\begin{lstlisting}[caption={Merging two queues},label={lst:concatenate-als},captionpos=b]
Node concatenateNodes(Node leftNode, Node rightNode, Node leftSubtreeMax) {
  if leftNode == nil:
    return rightNode
  else if rightNode == nil: 
    return leftNode
  else if leftNode.height == rightNode.height:
    Node result = Node(leftSubtreeMax, leftNode, rightNode)
    updateHeight(result)
    return result
  else if leftNode.height < rightNode.height:
    rightNode.left = concatenateNodes(leftNode, rightNode.left, leftSubtreeMax)
    updateHeight(rightNode)
    return rightNode
  else:
    leftNode.right = concatenateNodes(leftNode.right, rightNode, leftSubtreeMax)
    updateHeight(leftNode)
    return leftNode
		\end{lstlisting}

		First, we consider corner cases where one of the elements is $null$. This is needed to ensure correctness of the recursion. Then, if the left element is higher than the right, one step down is taken for the left element. If the right element is higher - we take a step down for it. If the heights are equal, the joining point is found and a new element must be created. At each step, it is necessary to update the height of current element because it changes.

		We begin analyzing the complexity of the $split$ procedure by determining the complexity of the $concatenateNodes$ procedure. At each iteration, a step is performed either to the left son of the current element or to the right one. The execution of the recursive procedure finishes by merging two elements. Since each step moves us down one level and a constant amount of work is performed for each level, the total complexity of the $concatenateNodes$ procedure is $O(h)=O(\log n)$.

		The $split$ procedure uses the $concatenateNodes$ function as a subroutine. The complexity of a $split$ call is equal to the complexity of $concatenateNodes$. The number of recursive $split$ calls for one such operation is $\log n$, so the total complexity of the procedure $\log^2 n$.

		The merge operation of two queues is reduced to the clamping of their root vertices by the $concatenateNodes$ procedure, so its complexity is $O(\log n)$.

\section{Algorithm analysis and performance evaluation}
\subsection{Complexity}
% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////
% ----------------------------------------------- COMPLEXITY ---------------------------------------------------------------------
% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

@@

			\begin{itemize}
				\item
				Preprocessing $O(n\log n)$.
				\item
				Recursive descent and splitting the set into 2 parts $O(1)$.
				\item
				Recursive ascent and merging parts of the convex hull $O(\log n)$.
				\begin{itemize}
					\item
					Transfer of the utmost points to upper parts of convex hulls $O(\log n)$.
					\item
					Finding the tangent for the upper parts of the hulls $O(\log n)$.
					\item
					Splitting and merging the upper parts $O(\log^2 n)$.
					\item
					Moving the utmost points to the bottom of the hulls $O(\log n)$.
					\item
					Finding the tangent for the upper parts of the hulls $O(\log n)$.
					\item
					Splitting and merging the upper parts $O(\log^2 n)$.
					\item
					Merging the lower parts of the hull $O(\log n)$.
					\item
					Normalization of the obtained lower part $O(\log n)$.
				\end{itemize}
			\end{itemize}

@@

@@

\subsection{Performance}
		A number of algorithm performance measurements were performed for different input sizes and the average number of recursive tasks per thread. The results are shown on the Fig. \ref{fig:performance}.

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.5\textwidth, height=0.3\textheight]{performance}}
			\caption{Performance data}
			\label{fig:performance}
		\end{figure}
\section{Conclusion}

@@

\begin{thebibliography}{00}
	\bibitem{ilastik} C. Sommer, C. Straehle, U. Köthe and F. A. Hamprecht, ``Ilastik: Interactive learning and segmentation toolkit,'' 2011 IEEE International Symposium on Biomedical Imaging: From Nano to Macro, Chicago, IL, 2011, pp.. 230-233.
	doi: 10.1109/ISBI.2011.5872394

	\bibitem{imaris} IMARIS Software. Accessed on: 15.04.2019. [Online]. Available: https://imaris.oxinst.com

	\bibitem{imagej} ImageJ: An open platform for scientific image analysis. Accessed on: 15.04.2019. [Online]. Available: https://imagej.net/Welcome

	\bibitem{aggarwal} А. Aggarwal et al., ``Parallel computational geometry,'' Algorithmica 3, 1-4 (November 1988), pp.. 293-327.

	\bibitem{atallah} M.J. Atallah, R. Cole, M.T. Goodrich, ``Cascading divide-and-conquer: a technique for designing parallel algorithms,'' SIAM J. Comput. 18, 1989, pp.. 499-532.

	\bibitem{cole} R. Cole, M.T. Goodrich, ``Optimal parallel algorithms for polygon and point-set problems,'' In Proceedings of the fourth annual symposium on Computational geometry (SCG '88), ACM, New York, NY, USA, pp.. 201-210.

	\bibitem{amato} N.M. Amato, M.T. Goodrich and E.A. Ramos, ``Parallel algorithms for higher-dimensional convex hulls,'' Proceedings 35th Annual Symposium on Foundations of Computer Science, Santa Fe, NM, USA, 1994, pp.. 683-694.

	\bibitem{chen} D.Z. Chen, ``Efficient geometric algorithms on the EREW PRAM,'' in IEEE Transactions on Parallel and Distributed Systems, vol. 6, no. 1, pp.. 41-47, Jan. 1995.
	doi: 10.1109/71.363412

	\bibitem{berkman} O. Berkman, B. Schieber, U. Vishkin, ``A fast parallel algorithm for finding the convex hull of a sorted point set,'' Internat. J. Comput. Geom. Appl. 6, 1996, pp.. 231-242.

	\bibitem{goodman} J.E. Goodman and J. O'Rourke, editors,  Handbook of Discrete and Computational Geometry, CRC Press LLC, Boca Raton, FL, 2004.

	\bibitem{akl} Selim G. Akl and Kelly A. Lyons, Parallel Computational Geometry, Englewood Cliffs: Prentice-Hall, 1993.

	\bibitem{jaja} J. JaJa, An Introduction to Parallel Algorithms, Amsterdam: Addison-Wesley, 1992.

	\bibitem{leeuwen} J. Van Leeuwen, Handbook of Theoretical Computer Science, Amsterdam: Elsevier/The MIT Press, 1990.

	\bibitem{reif} J.H. Reif, Synthesis of Parallel Algorithms, San Mateo: Morgan Kaufmann, 1993.

	\bibitem{tereshchenko} V. N. Tereshchenko, A. V. Anisimov, Recursion and parallel algorithms in geometric modeling problems, Journal: Cybernetics and Systems Analysis: Vol. 46, N 2 (2010), pp.. 173-184.

	\bibitem{overmars} Mark H. Overmars, Jan van Leeuwen, ``Maintenance of configurations in the plane,'' Journal of Computer and System Sciences, Volume 23, Issue 2, 1981, pp.. 166-204.

	\bibitem{aho} Alfred V. Aho and John E. Hopcroft, The Design and Analysis of Computer Algorithms, 1st ed. Addison-Wesley Longman Publishing Co., Inc., Boston, MA, USA, 1974.
\end{thebibliography}
\end{document}

>>>>

\documentclass[a4paper,UKenglish,cleveref, autoref]{socg-lipics-v2019}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"
%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\lstset{
	language=Java,
	breaklines,
}

\title{Unified data structures for solving optimally a set of interrelated computational geometry problems} 
\titlerunning{Data structures for solving a set of computational geometry problems}%optional, please use if title is longer than one line

\author{Vasyl Tereshchenko}{
	Faculty of computer science and cybernetics, Taras Shevchenko National University of Kyiv, Ukraine}
{vtereshch@gmail.com}{}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional
%\footnote{}
\author{Semen Chudakov}{Faculty of computer science and cybernetics, Taras Shevchenko National University of Kyiv, Ukraine}
{semen.chudakov7@gmail.com}{}{}

\authorrunning{V. Tereshchenko and S. Chudakov}%TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Vasyl Tereshchenko and Semen Chudakov}%TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[500]{Theory of computation~Computational geometry}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Algorithmic tools,   Computational geometry, Interrelated problems set, Unified algorithmic environment, Concatenable queue}%TODO mandatory; please add comma-separated list of keywords

\category{}%optional, e.g. invited paper

\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{}%optional

%\nolinenumbers %uncomment to disable line numbering

%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{}
\EventNoEds{2}
\EventLongTitle{36th International Symposium on Computational Geometry (SoCG 2020)}
\EventShortTitle{SoCG 2020}
\EventAcronym{SoCG}
\EventYear{2020}
\EventDate{June  23--26, 2020}
\EventLocation{Zürich, Switzerland}
\EventLogo{socg-logo}
\SeriesVolume{}
\ArticleNo{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
	The paper is devoted to the development and  of an efficient algorithmic model for solving a set of interrelated computational geometry problems. To do this, a unified algorithmic environment with unified data structures is created, which allows to implement complex use cases efficiently with respect to computational resources. We build the environment on the basis of the ``divide and conquer'' strategy. 
	Once a convex hull is key to a set of computational geometry problems, we offer a concatenable queue data structure to maintain it. The data structure is implemented in a form of a binary tree. This allows to perform operations needed in algorithm for a set of problems in $O(\log n)$ time. Furthermore we offer a way to execute the algorithms both sequentially and in parallel.
	In the future the algorithmic environment can be improved to support other computational models with similar properties for solving problems. As an example, the Voronoi diagram or the Delaunay triangulation can be considered.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

@@

\section{Unified algorithmic environment}
\label{sec:unified-algorithmic-environment}
\subsection{Algorithms stages}

@@

	\begin{equation}
	\lambda(a_1, a_2)=\frac{y(a_2)-y(a_1)}{x(a_2)-x(a_1)}
	\end{equation}

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.8\textwidth, height=0.3\textheight]{ch_union}
		\caption{Merging step of two hulls}
		\label{fig:ch_union}
	\end{figure}

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.8\textwidth, height=0.3\textheight]{incorect_search}
		\caption{Example of an incorrect position of the utmost left in the left sub-hull}
		\label{fig:incorect_search}
	\end{figure}

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.8\textwidth, height=0.3\textheight]{incorect_edge_points}
		\caption{Example of a convex hull for a wrong position of the utmost left points of the left sub-hull}
		\label{fig:incorect_edge_points}
	\end{figure}

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.8\textwidth, height=0.3\textheight]{incorect_lower_subhull}
		\caption{An example of a non-integral hull after merging along the reference lines}
		\label{fig:incorect_lower_subhull}
	\end{figure}

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.8\textwidth, height=0.3\textheight]{correct_convex_hull}
		\caption{Correctly constructed convex hull}
		\label{fig:correct_convex_hull}
	\end{figure}

@@

	\begin{lstlisting}[caption={Algorithm model based on the ``divide-and-conquer'' principle},label={lst:algmodel},captionpos=t,float,abovecaptionskip=-\medskipamount]
interface DaCAlgorithm[IT, OT]:
    boolean isBaseCase(IT input)
    int inputSize(IT input)

@@

    OT merge(OT first, OT second)
    Pair[IT, IT] divide(IT input)
	\end{lstlisting}

\section{Algorithm analysis and performance evaluation}
\label{sec:algorithm-analysis-and-performance-evaluation}
\subsection{Complexity}

% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////
% ----------------------------------------------- COMPLEXITY ---------------------------------------------------------------------
% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

@@

		\begin{enumerate}
			\item
			Preprocessing $O(n\log n)$.
			\item
			Recursive descent and splitting the set into 2 parts $O(1)$.
			\item
			Recursive ascent and merging parts of the convex hull $O(\log n)$.
			\begin{alphaenumerate}
				\item
				Transfer of the utmost points to upper parts of convex hulls $O(\log n)$.
				\item
				Finding the tangent for the upper parts of the hulls $O(\log n)$.
				\item
				Splitting and merging the upper parts $O(\log^2 n)$.
				\item
				Moving the utmost points to the bottom of the hulls $O(\log n)$.
				\item
				Finding the tangent for the upper parts of the hulls $O(\log n)$.
				\item
				Splitting and merging the upper parts $O(\log^2 n)$.
				\item
				Merging the lower parts of the hull $O(\log n)$.
				\item
				Normalization of the obtained lower part $O(\log n)$.
			\end{alphaenumerate}
		\end{enumerate}

@@

\subsection{Performance}
\label{sec:performance}

@@

	A number of algorithm performance measurements were performed for different input sizes and the average number of recursive subproblems per thread. The results are shown on the Fig. \ref{fig:performance}.

	\begin{figure}[h]
		\centering
		\includegraphics[width=1\textwidth, height=0.4\textheight]{performance}
		\caption{Performance data}
		\label{fig:performance}
	\end{figure}

\section{Conclusion}

@@

\bibliography{article-SoCG-EA}
\end{document}

### 2019-12-23 SoCG/article-SoCG.tex
SoCG EA, SoCG, acm siggraph + restructuring

  OT merge(OT first, OT second)
  Pair[IT, IT] divide(IT input)

		\end{lstlisting}

@@

\section{Implementation details}
\subsection{Concatenable queue}

@@

  updateHeight(result)
  return result
			\end{lstlisting}
		\end{figure}

@@

\section{Algorithm analysis and performance evaluation}
\subsection{Complexity}
% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////
% ----------------------------------------------- COMPLEXITY ---------------------------------------------------------------------
% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

@@

@@

\subsection{Performance}
		A number of algorithm performance measurements were performed for different input sizes and the average number of recursive tasks per thread. The results are shown on the Fig. \ref{fig:performance}.

		\begin{figure}[htbp]
			\centerline{\includegraphics[width=0.5\textwidth, height=0.3\textheight]{performance}}
			\caption{Performance data}
			\label{fig:performance}
		\end{figure}
\section{Conclusion}

@@

\begin{thebibliography}{00}
	\bibitem{ilastik} C. Sommer, C. Straehle, U. Köthe and F. A. Hamprecht, ``Ilastik: Interactive learning and segmentation toolkit,'' 2011 IEEE International Symposium on Biomedical Imaging: From Nano to Macro, Chicago, IL, 2011, pp.. 230-233.
	doi: 10.1109/ISBI.2011.5872394

	\bibitem{imaris} IMARIS Software. Accessed on: 15.04.2019. [Online]. Available: https://imaris.oxinst.com

	\bibitem{imagej} ImageJ: An open platform for scientific image analysis. Accessed on: 15.04.2019. [Online]. Available: https://imagej.net/Welcome

	\bibitem{aggarwal} А. Aggarwal et al., ``Parallel computational geometry,'' Algorithmica 3, 1-4 (November 1988), pp.. 293-327.

	\bibitem{atallah} M.J. Atallah, R. Cole, M.T. Goodrich, ``Cascading divide-and-conquer: a technique for designing parallel algorithms,'' SIAM J. Comput. 18, 1989, pp.. 499-532.

	\bibitem{cole} R. Cole, M.T. Goodrich, ``Optimal parallel algorithms for polygon and point-set problems,'' In Proceedings of the fourth annual symposium on Computational geometry (SCG '88), ACM, New York, NY, USA, pp.. 201-210.

	\bibitem{amato} N.M. Amato, M.T. Goodrich and E.A. Ramos, ``Parallel algorithms for higher-dimensional convex hulls,'' Proceedings 35th Annual Symposium on Foundations of Computer Science, Santa Fe, NM, USA, 1994, pp.. 683-694.

	\bibitem{chen} D.Z. Chen, ``Efficient geometric algorithms on the EREW PRAM,'' in IEEE Transactions on Parallel and Distributed Systems, vol. 6, no. 1, pp.. 41-47, Jan. 1995.
	doi: 10.1109/71.363412

	\bibitem{berkman} O. Berkman, B. Schieber, U. Vishkin, ``A fast parallel algorithm for finding the convex hull of a sorted point set,'' Internat. J. Comput. Geom. Appl. 6, 1996, pp.. 231-242.

	\bibitem{goodman} J.E. Goodman and J. O'Rourke, editors,  Handbook of Discrete and Computational Geometry, CRC Press LLC, Boca Raton, FL, 2004.

	\bibitem{akl} Selim G. Akl and Kelly A. Lyons, Parallel Computational Geometry, Englewood Cliffs: Prentice-Hall, 1993.

	\bibitem{jaja} J. JaJa, An Introduction to Parallel Algorithms, Amsterdam: Addison-Wesley, 1992.

	\bibitem{leeuwen} J. Van Leeuwen, Handbook of Theoretical Computer Science, Amsterdam: Elsevier/The MIT Press, 1990.

	\bibitem{reif} J.H. Reif, Synthesis of Parallel Algorithms, San Mateo: Morgan Kaufmann, 1993.

	\bibitem{tereshchenko} V. N. Tereshchenko, A. V. Anisimov, Recursion and parallel algorithms in geometric modeling problems, Journal: Cybernetics and Systems Analysis: Vol. 46, N 2 (2010), pp.. 173-184.

	\bibitem{overmars} Mark H. Overmars, Jan van Leeuwen, ``Maintenance of configurations in the plane,'' Journal of Computer and System Sciences, Volume 23, Issue 2, 1981, pp.. 166-204.

	\bibitem{aho} Alfred V. Aho and John E. Hopcroft, The Design and Analysis of Computer Algorithms, 1st ed. Addison-Wesley Longman Publishing Co., Inc., Boston, MA, USA, 1974.
\end{thebibliography}
\end{document}

>>>>

    OT merge(OT first, OT second)
    Pair[IT, IT] divide(IT input)
	\end{lstlisting}

@@

\section{Implementation details}
\label{sec:implementation-details}
\subsection{Concatenable queue}

@@

    updateHeight(result)
    return result

		\end{lstlisting}
	\end{figure}

@@

\section{Algorithm analysis and performance evaluation}
\label{sec:algorithm-analysis-and-performance-evaluation}
\subsection{Complexity}

% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////
% ----------------------------------------------- COMPLEXITY ---------------------------------------------------------------------
% //////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

@@

\subsection{Performance}

@@

	A number of algorithm performance measurements were performed for different input sizes and the average number of recursive subproblems per thread. The results are shown on the Fig. \ref{fig:performance}.

	\begin{figure}[h]
		\centering
		\includegraphics[width=1\textwidth, height=0.4\textheight]{performance}
		\caption{Performance data}
		\label{fig:performance}
	\end{figure}

\section{Conclusion}
\label{sec:coclusion}

@@

\bibliography{article-SoCG}
\end{document}
